---
title: "ThemaOpdracht09Log"
author: "Wisse Schuuring"
date: "9/14/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Day 1 Tuesday 14th of September, 2021

From the github of michielnoback chosen the "Crab body metrics" dataset.
Downloaded the dataset from https://www.kaggle.com/inputblackboxoutput/crab-body-metrics

Description
This dataset has 200 rows and 8 columns, describing 5 morphological measurements on 50 crabs each of two colour forms and both sexes, of the species Leptograpsus variegatus collected at Fremantle, W. Australia.

SP = Species: "B" or "O" for blue or orange.
sex = M- Male F - Female
index = Unique row identifier
FL = Frontal lobe size (mm)
RW = Rear width (mm)
CL = Carapace length (mm)
CW = Carapace width (mm)
BD = Body depth (mm)

Source
Campbell, N.A. and Mahon, R.J. (1974) A multivariate study of variation in two species of rock crab of genus Leptograpsus. Australian Journal of Zoology 22, 417–425.

References
Venables, W. N. and Ripley, B. D. (2002) Modern Applied Statistics with S. Fourth edition. Springer.

Created a Github account along with a repo, and pushed the dataset and log file along with everything else "git" into it.

Read the csv file and attempted to create a handful of plots with its contents.

Succesfully plotted the crabs' Body Depth in relation to its Rear Width, and comparing it with its species and sex.

```{r}
CrabData <- read.csv(file = 'data/CrabData.csv')
CrabData

ggplot(CrabData, aes(y = BD, x = RW, color=sp, shape = sex)) +
  geom_point() +
  scale_x_continuous() +
  geom_jitter() +
  scale_color_manual(values=c("Blue","Orange"))



```