---
title: "ThemaOpdracht09Log"
author: "Wisse Schuuring"
date: "9/14/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Day 1 Tuesday 14th of September, 2021

From the github of michielnoback chosen the "Crab body metrics" dataset.
Downloaded the dataset from https://www.kaggle.com/inputblackboxoutput/crab-body-metrics

Description
This dataset has 200 rows and 8 columns, describing 5 morphological measurements on 50 crabs each of two colour forms and both sexes, of the species Leptograpsus variegatus collected at Fremantle, W. Australia.

SP = Species: "B" or "O" for blue or orange.
sex = M- Male F - Female
index = Unique row identifier
FL = Frontal lobe size (mm)
RW = Rear width (mm)
CL = Carapace length (mm)
CW = Carapace width (mm)
BD = Body depth (mm)

Source
Campbell, N.A. and Mahon, R.J. (1974) A multivariate study of variation in two species of rock crab of genus Leptograpsus. Australian Journal of Zoology 22, 417–425.

References
Venables, W. N. and Ripley, B. D. (2002) Modern Applied Statistics with S. Fourth edition. Springer.

Created a Github account along with a repo, and pushed the dataset and log file along with everything else "git" into it.

Read the csv file and attempted to create a handful of plots with its contents.

Successfully plotted the crabs' Body Depth in relation to its Rear Width, and comparing it with its species and sex.

```{r}
#Read in the libraries
library(ggplot2)

#Read in the dataset
CrabData <- read.csv(file = 'data/CrabData.csv')
CrabData 

#Give the 5 values of the dataset
summary(CrabData)

#Create a plot of both species
ggplot(CrabData, aes(y = RW, x = BD, color=sex, shape = sp)) +
  geom_point() +
  scale_x_continuous() +
  geom_jitter() +
  geom_smooth() +
  scale_color_manual(values=c("Blue","Orange"))


#Create a boxplot
boxplot(CrabData$RW, CrabData$BD)

```



# Day 2 Wednesday 15th of September, 2021



Successfully created two separate plots for the two species.

```{r}
#Read in the libraries
library(ggplot2)
library(dplyr)

#Read in the dataset
CrabData <- read.csv(file = 'data/CrabData.csv')
CrabData 

#Filter the datasets on their species
CrabDataB <- filter(CrabData, sp == "B")
CrabDataO <- filter(CrabData, sp == "O")

#Give the 5 values of the dataset
summary(CrabData)

#Create a plot of both species
ggplot(CrabData, aes(y = RW, x = BD, color=sex, shape = sp)) +
  geom_point() +
  scale_x_continuous() +
  geom_jitter() +
  geom_smooth() +
  scale_color_manual(values=c("Blue","Orange"))

#Create a plot of only the blue species 
ggplot(CrabDataB, aes(y = RW, x = BD)) +
  geom_point() +
  scale_x_continuous() +
  geom_jitter() +
  geom_smooth(col = "blue")

#Create a plot of only the orange species
ggplot(CrabDataO, aes(y = RW, x = BD)) +
  geom_point() +
  scale_x_continuous() +
  geom_jitter() +
  geom_smooth(col = "orange")

#Create a boxplot
boxplot(CrabData$RW, CrabData$BD)

```
The first canonical variate, which differentiates between the two species, represents a contrast
between the carapace width relative to the width of the front lip and the depth of the body;
the blue-form species has a greater relative carapace width than has the orange form. 

```{r}
#Create a plot showing the carapace width against the frontal lobe size of the two species
ggplot(CrabData, aes(y=CW, x=FL, col=sp)) +
  geom_point() +
  geom_jitter() +
  scale_colour_manual(values = c("Blue","Orange"))

```
The second canonical variate, which presents a contrast between the rear width and the carapace
length, identifies males and females within each species; males have a greater relative carapace
length than have females. 

```{r}
ggplot(CrabData, aes(y=RW, x=CL, col=sex)) +
  geom_point() +
  geom_jitter() +
  scale_colour_manual(values = c("Red","Dark Green"))
```

All individuals, including 20 not used in the study, were correctly identified for colour form.
The clear separation of the blue and orange forms achieved by canonical analysis supports the
previously determined specific status of the two forms. 

```{r}
str(CrabData)
stinky

```